---
title: "Occupancy Status"
description: |
  A short description of the post.
date: 2021-11-01
output:
  distill::distill_article:
    self_contained: false
draft: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Intro

In my last post, I looked at Census data for how many housing units there were in the Memphis area. According to the 2015-2019 5-Year ACS, there are 567,073 housing units in the Memphis metro area; 405,184 within Shelby County; and 299,913 within the city of Memphis.

This post will look at the occupancy status of housing units: whether they are occupied or vacant.

# Metro Area

```{r}
library(tidyverse)
library(tidycensus)
library(sf)
```

```{r include=FALSE}
occStat <- get_acs(
  geography = "cbsa",
  variables = c(
    Occupied = "B25002_002",
    Vacant = "B25002_003"
  ),
  summary_var = "B25002_001",
  cache_table = TRUE
) %>%
  slice_max(summary_est, n = 100)
```

```{r}
occStatPct <- occStat %>% 
  mutate(percent = 100 * (estimate / summary_est)) %>%
  select(NAME, variable, estimate, percent)
```

In my previous post, I made a table of the top 50 metro areas in the US by number of housing units, where Memphis ranked 43rd.

Below is a table ranking those metros by vacancy.

```{r}
library(rmarkdown)
```

```{r}
vacMetro <- occStatPct %>% 
  filter(variable == "Vacant") %>% 
  select(-variable) %>% 
  arrange(desc(percent)) %>% 
  mutate(NAME = str_remove(NAME, "-.*$")) %>% 
  mutate(NAME = str_remove(NAME, ",.*$")) 
paged_table(vacMetro)
```

Unfortunately, Memphis ranks high on this list, with a vacancy rate of 12.32%. Of the top 50 metros (by number of housing units), Memphis has the 10th highest vacancy rate. According to the 2015-2019 5-Year ACS, there are 69,841 vacant housing units in the Memphis metro area.

```{r include=FALSE}
cbsa <- get_acs(
  geography = "cbsa",
  variables = "B25002_001",
  geometry = TRUE,
  cache_table = TRUE
) %>% filter(str_detect(NAME, "Memphis"))

memST <- list("TN", "MS", "AR")
```

```{r include=FALSE}
occCo <- map_dfr((memST), ~ {
  get_acs(
    geography = "county",
    state = .x,
    variables = c(
    Occupied = "B25002_002",
    Vacant = "B25002_003"),
  summary_var = "B25002_001",
    geometry = TRUE,
    cache_table = TRUE
  )
}) %>% 
  st_filter(cbsa, .predicate = st_within) %>% 
  mutate(area = as.numeric(st_area(geometry)),
         percent = 100 * (estimate / summary_est))
```

```{r}
library(ggiraph)
library(patchwork)
```

```{r}
vacCo <- occCo %>% filter(variable == "Vacant")
```

```{r}
vacCoMap <- ggplot(vacCo, aes(fill = percent)) +
  geom_sf() +
  theme_void()

vacCoMap
```
