---
title: "Income & Affordable Housing Table"
description: |
  A table showing the maximum amount a household can spend on housing without being cost burdened.
author:
  - name: Sarah Johnson
date: 2022-06-19
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The federal government determines **affordable housing** to be no more than 30% of a person's income. Someone paying more than 30% is considered **cost burdened**; more than 50% is **extremely cost burdened**.

The following table lists incomes and corresponding affordable rents. This table does not account for any taxes that decrease take-home wages (and thus lowering how much rent one can afford) or if someone with an hourly wage works more/less than 40 hours a week.

**Yearly Income** is hourly wage \* 40 hours a week \* 52 weeks a year.

**Monthly Housing** is 30% of the yearly wage, divided by 12 months of the year.

<br>

```{r library}
library(tidyverse)
library(DT)
income <- read_csv("income.csv")
```

```{r affRents}
affRents <- income %>% 
  mutate(wageYear = wageHour * 40 * 52,
         rent = wageYear * .3 / 12)
```

```{r}
datatable(affRents,
          rownames = FALSE,
          colnames = c("Hourly Wage", "Yearly Income", "Monthly Housing"),
          caption = "Table 1: Incomes and maximum affordable housing costs",
          options = list(pageLength = 20)
          ) %>% 
  formatCurrency(1) %>% 
  formatCurrency(2:3, digits = 0)
```

