---
title: "tracts-timeseries"
description: |
  A short description of the post.
date: 2021-10-20
output:
  distill::distill_article:
    self_contained: false
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Boundaries Over Time

As places grow and change, so do their boundaries. It's important we know how boundaries have changed over time before we began analyzing time-series data.

We'll need some extra packages for this section to create our final maps.

```{r}
library(tidyverse)
library(tigris)
library(patchwork)
library(glue)
```

## Tracts within County

To create a map over a certain period of time, use `map(seq())` functions. Set your start year, end year, and interval, and define your **tigris** geography, setting `year = .x`. In the same mapping function, set up your `ggplot()` and use the `glue()` function to automatically label maps with the year and number of rows for that year (equal to the number of tracts).

```{r}
tractsPlot <- map(seq(1990, 2020, 10), ~{
  tractsYear <- tracts("TN", "Shelby", year = .x, cb = TRUE)
  
  ggplot(tractsYear) +
    geom_sf() +
    theme_void() +
    labs(title = glue("{.x}: {nrow(tractsYear)} tracts"))
})
```

Then we can use **patchwork** to create up our plots.

```{r}
(tractsPlot[[1]] + tractsPlot[[2]]) / 
  (tractsPlot[[3]] + tractsPlot[[4]])
```

Shelby County gained 65 tracts from 1990 to 2020.
