---
title: "CE1-5 Recap"
description: |
  A summary of my progress wrangling Memphis code enforcement data.
author:
  - name: Sarah Johnson
date: 2022-02-19
output:
  distill::distill_article:
    self_contained: false
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      code_folding=TRUE)
```

## Purpose

When a renter encounters a home maintenance problem, they rely on their landlord to complete the repairs. In Tennessee, landlords who do not complete repairs in a timely manner are in breach of the Uniform Residential Landlord-Tenant Act (URLTA). Renters who wish to remain in their home and seek remedies under URLTA are expected to report problems to their city's code enforcement (CE). However, tenants may choose not to report problems due to fears of retaliation; while URLTA is supposed to protect against retaliation, landlords are also allowed to evict tenants without cause.

For my capstone, I wanted to analyze Memphis's code enforcement dataset to answer a few key questions:

-   How often do renters report problems to code enforcement?

-   What are the most common problems renters report?

-   How are these requests resolved?

-   How long does the process take?

Memphis code enforcement data is publicly available through the Memphis Data Hub. The dataset includes all 311 service requests since 2016, it is updated daily, and contains over 1.3 million rows at the time of this posting.

However, despite the vastness of the dataset, there was way no way to easily answer my questions. There is no way to filter for structural problems, much less renter-related problems. Many rental-related problems are categorized under "Miscellaneous", and useful, relevant information is primarily contained in a single, manually-entered, catch-all column that makes data analysis difficult.

The purpose of this CE series has been to develop a method to extract useful information from this dataset to find the answers to my questions.

## Introducing the Dataset

For my analysis, I have been using a version of the dataset downloaded on December 3, 2021 that contains 1,311,942 rows and 53 columns. Despite the large size of the dataset, only a portion of the information is relevant to my research.

Every 311 call made to the city opens a service request (SR), and each row of the dataset is a unique SR. Not all SRs are related to code enforcement (for instance, citizens can report potholes to 311 but they are not under CE's purview). After filtering for CE, there are 165,836 rows that remain.

In [CE1](https://sarahjohnson.io/posts/2021-12-03-311-data/) I walk through all columns included in the dataset, briefly explaining the purpose of each, which contain duplicate information, and which are most useful to my research. Most of the columns are related to location or date data and only a handful contain more pertinent information. The below table lists the columns I most frequently use, how I abbreviate them in my code, and a brief summary of their purpose. I also use the `DEPARTMENT` category at the beginning of each post to filter the data for code enforcement requests.

| Column Name             | Abbr. | Summary                                                                                                                                                |
|------------------|------------------|-------------------------------------|
| INCIDENT_NUMBER         | sr    | A unique number for each service request; the dataset's key.                                                                                           |
| REQUEST_TYPE            | \-    | An SR can be classified under one of 14 request types. Rental-related SRs are categorized under "Substandard, Derelict Struc" or "Code Miscellaneous". |
| RESOLUTION_CODE         | rc    | A category for inspectors to select how the SR was resolved. This column is abbreviated.                                                               |
| RESOLUTION_CODE_MEANING | rcm   | Contains the same information as the above column, but in non-abbreviated form.                                                                        |
| RESOLUTION_SUMMARY      | rs    | A catch-all column to contain any notes and updates related to the SR.                                                                                 |

Of the 53 columns in the dataset, only the `RESOLUTION_CODE` and the `RESOLUTION_SUMMARY` provide particularly useful qualitative information for each service request. The lack of a `REQUEST_TYPE` specifically geared towards rental requests makes analysis particularly hard, because these SRs are often categorized under "Miscellaneous" with no easy way to identify if they are related to structural problems.

While the `RESOLUTION_SUMMARY` contains a lot of valuable information, extracting information for analysis has proven difficult. Because all summaries are manually entered, there are inconsistent data entry methods and frequent typos. Still, there are repeated patterns to how inspectors enter information that makes some data extraction possible.

## The Goal: Tidy Data

My overall goal is to convert Memphis's code enforcement data into a tidy dataset. **Tidy data** is a standardized way to organize data that makes it easy for analysts to quickly find the information they're looking for. In tidy data:

1\. Each variable has its own column.

2\. Each observation has its own row.

3\. Each value has its own cell.

![Three rules make a dataset tidy: variables are in columns, observations are in rows, and values are in cells ([Wickham & Grolemund](https://r4ds.had.co.nz/tidy-data.html#introduction-6)).](https://d33wubrfki0l68.cloudfront.net/6f1ddb544fc5c69a2478e444ab8112fb0eea23f8/91adc/images/tidy-1.png)

To make the dataset tidy, I plan to:

-   Eliminate the need for the `RESOLUTION_SUMMARY` column

-   Collapse **date** and **location** columns

-   Recode the `RESOLUTION_CODE` column to be more consistent

-   Create new categories under `REQUEST_TYPE` for structural violations

These goals often overlap. Dates, locations, and resolution codes are frequently mentioned in the `RESOLUTION_SUMMARY` column. Sometimes the only way to determine whether a request is related to a rental unit is to search that column for keywords like "tenant". Cleaning up this column and extracting useful information is the largest obstacle to tidying the dataset.

## Progress on Resolution Summary

In CE3-5, I have worked to develop a way to extract data from the `RESOLUTION_SUMMARY` column. To do this, I have created a pattern in my code:

1.  Identify the information I wish to extract.
2.  Extract the data from the `RESOLUTION_SUMMARY` into a new column.
3.  Remove the information from the `RESOLUTION_SUMMARY`.
4.  Filter the dataset to only show rows that still contain information in the `RESOLUTION_SUMMARY`.

I focused on numerical data in [CE3](https://sarahjohnson.io/posts/2022-01-10-311-data-3/), including dates, references to other service requests, and phone numbers. In [CE4](https://sarahjohnson.io/posts/2022-01-11-311-data-4/) I identified which resolution summaries duplicated the `RESOLUTION_CODE` field, and created a list of additional ways service requests were resolved that were not previously coded. In [CE5](https://sarahjohnson.io/posts/2022-01-22-311-data-5/) I noted that inspectors would often sign off on service requests by entering their name, so I created a list of inspectors mentioned in the dataset. In that post I also created a new column to contain updates to requests, similar to the resolution codes, but not resolutions.

## The Code

First, import the dataset. I alter some columns to all lowercase to make data matching easier.

```{r}
library(tidyverse)

zService_Requests_since_2016 <-
  read_csv("../_data/Service_Requests_since_2016.csv") %>% 
  mutate(RESOLUTION_SUMMARY = str_to_lower(RESOLUTION_SUMMARY),
         RESOLUTION_CODE = str_to_lower(RESOLUTION_CODE),
         RESOLUTION_CODE_MEANING = str_to_lower(RESOLUTION_CODE_MEANING))
```

Filter the `DEPARTMENT` for code enforcement data and the `REQUEST_TYPE` for "derelict" and "miscellaneous", the two request types relevant to rental housing. I further narrowed my focus to four columns: the `INCIDENT_NUMBER`, or service request (SR) number, to use as a key, and three columns related to the request's resolution.

```{r}
zCE <- zService_Requests_since_2016 %>%
  filter(DEPARTMENT == "Code Enforcement" &
           str_detect(REQUEST_TYPE, "Misc|Derelict")) %>%
  select(sr = INCIDENT_NUMBER,
         rc = RESOLUTION_CODE,
         rcm = RESOLUTION_CODE_MEANING,
         rs = RESOLUTION_SUMMARY) %>%
  filter(!is.na(rs))
```

### Extract numbers

Using the code from [CE3](https://sarahjohnson.io/posts/2022-01-10-311-data-3/), separate numerical data into new columns. The following code will be used to find mentions of service request numbers, dates, and tax sale numbers.

```{r}
vSR <- c(
  "\\b(sr)?\\d{7}\\b",
  "(please )?(see|(ref(er)?(ence)?)\\b( to)?)?( )?(\\b(see)?s[/ \\.]*r\\b|service request|ticket)( )?(#)?( )?(\\.)?",
  "(please )?(see|(ref(er)?(ence)?)\\b( to)?)( )?(#)?( )?(\\.)?"
) %>% 
  str_c(collapse = "|")

vMonths <-
  c("january", "february", "march", "april", 
    "may", "june", "july", "august", 
    "september", "october", "november", "december",
    str_c("\\b",
      c("jan", "feb", "mar", "apr", "jun", "jul", 
        "aug", "sept", "oct", "nov", "dec"),
      "(\\.)?\\b"
      )
    )

vDate <- c(
  "\\b\\d{1,2}(:\\d{2})?(:\\d{2})?( )?(a(\\.)?m(\\.)?|p(\\.)?m(\\.)?)",
  str_c(vMonths, "( )?", "\\d{1,2}", "(nd|th|rd)?", "((,)?( )?\\d{2,4})?"),
  str_c("\\d{1,2}", "(nd|th|rd)? ", vMonths, "((,)?( )?\\d{2,4})?"),
  str_c("\\d{1,2}", "\\W", vMonths, "\\W", "\\d{2,4}"),
  str_c("\\d{1,2}", "(nd|th|rd)?", " of ", vMonths),
  "(?<=(\\b|\\D))\\d{1,2}( )?\\W{1,2}\\d{1,2}( )?\\W( )?\\d{2}(?=\\b|\\D)",
  "(?<=(\\b|\\D))\\d{1,2}( )?\\W\\d{1,2}\\W( )?\\d{4}",
  "\\b\\d{1,2}\\W{1,2}\\d{1,2}\\b",
  "\\b\\d{1,2}\\W\\d{4}\\b"
  ) %>%
  str_c(collapse = "|")

vTax <- "(?<=tax sale( (number|2019))? )\\d{2}(\\.)?\\d{2}(?!(\\W)?\\d)"
```

I am hiding the code used to identify phone numbers from these blog posts due to privacy concerns.

```{r include=FALSE}
vPH <- "(ph)?[# ]*(\\d{3}\\W)?\\b\\d{3}\\b\\W\\d{4}"
```

Use the above code to extract this information from the resolution summary into new columns. I also added a section of code (containing `"\\W"`) to remove non-alphanumeric characters, due to punctuation and symbols complicating future code.

```{r}
zNumbers <- zCE %>%
  mutate(
    #' create new columns & remove extracted data from rs
    ref = str_extract_all(rs, "\\b(sr)?\\d{7}\\b"),
    rs = case_when(
      str_detect(rs, "\\b(sr)?\\d{7}\\b") ~ str_replace_all(rs, vSR, " "),
      TRUE ~ rs
    ),
    taxSale = str_extract(rs, vTax),
    rs = str_remove(rs, vTax),
    date = str_extract_all(rs, vDate),
    rs = str_remove_all(rs, vDate),
    rs = str_remove_all(rs, vPH),
    #' clean up rs, rc, rcm
    rs = str_replace_all(rs, "\\W", " "),
    rc = str_replace_all(rc, "\\W", " "),
    rcm = str_replace_all(rcm, "\\W", " "),
    rs = str_squish(rs),
    rc = str_squish(rc),
    rcm = str_squish(rcm),
    rs = na_if(rs, "")
  )
```

I remove rows where the resolution summary and resolution code columns are empty, to see which rows still need to be cleaned up.

```{r}
zNum <- zNumbers %>% 
  filter(!is.na(rs) & !is.na(rc))

zNum %>% count(rs) %>% arrange(desc(n))
```

There are 45,333 rows that still contain information in the resolution summary or resolution code.

### Extract Resolution Codes

Using the code from [CE4](https://sarahjohnson.io/posts/2022-01-11-311-data-4/), identify instances where the resolution summary contains duplicate information found in the resolution code. I also created an incomplete list of other resolutions commonly mentioned but not included in the resolution code column.

```{r, code_folding=TRUE}
vErr <- str_c("^", c("shelby", "board and secure", "hold", "other"), "$") %>% str_c(collapse = "|")

zRCs <- zNum %>% 
  select(rc, rcm) %>% 
  pivot_longer(cols = c(rcm, rc), names_to = "remo", values_to = "rcs") %>% 
  select(rcs) %>% 
  mutate(rcs = str_remove(rcs, vErr),
         rcs = na_if(rcs, "")) %>%
  filter(!is.na(rcs)) %>% 
  distinct()

vRCs <- zRCs$rcs

vRCm <- c(
  "((shelby )?county|city).*property",
  "scop",
  "justified active",
  "dismissed",
  "(complaint.*)?not justified",
  "no cooperation",
  "jw(dm|rh)",
  "cvob", "fm11",
  "vobr", "vobd", "voar(s)?",
  "rvso",
  "npfm11",
  "v( )?o",
  "rdw",
  "noh3",
  "^closed( violation)?( out)?",
  "^hold$",
  "close other"
  )

vRC <- str_c("\\b", c(vRCs, vRCm), "\\b") %>% str_c(collapse = "|")
```

I extracted this information into a new column, `rc2`, a temporary secondary resolution code column.

```{r}
vInspectors <- str_c("\\b", c(
  "(c|cheryl)?( )?clausel",
  "(q|quatarious)?( )?gilc(l)?hrest",
  "(r|robert)?( )?gray",
  "(j|judith)?( )?norman",
  "[amw]?( )?lewis( nickson)?",
  "[arc]?( )?hill",
  "[dr]?( )?butler",
  "[ej]?( )?cooper",
  "[al]?( )?davis",
  "[ja]*( )?irwin",
  "[ce]?( )?jones",
  "[rn]?( )?rashada",
  "[rd]?( )?witter",
  "[ac]?( )?wi[li]*am(s)?",
  "(e)?( )?blow",
  "(k)?( )?boone",
  "(c)?( )?boykins",
  "(b)?( )?bratten",
  "(r)?( )?brown",
  "(c)?( )?bryant",
  "(s)?( )?burgess",
  "(d)?( )?burton",
  "(l)?( )?busby",
  "(t)?( )?cobb",
  "(p)?( )?cook",
  "(a)?( )?collins",
  "(s)?( )?echols",
  "(v)?( )?ervin",
  "(t)?( )?fry",
  "(s)?( )?gaspar",
  "(j)?( )?golden",
  "(d)?( )?grafton",
  "(c)?( )?higgenbottom",
  "(s)?( )?howell",
  "(j)?( )?hymon",
  "(j)?( )?jackson",
  "(m)?( )?james",
  "(r)?( )?lake",
  "(d)?( )?massey",
  "(m)?( )?middlebrooks",
  "(s)?( )?milan",
  "(d)?( )?mi(t)?chell",
  "(j)?( )?mos[lbe]*y",
  "(v)?( )?moses",
  "(t)?( )?neff",
  "(r)?( )?nelson",
  "(t)?( )?norwood",
  "(b)?( )?pa(t)+erson", 
  "(k)?( )?payne",
  "(r)?( )?peete(e)?",
  "(t)?( )?pegues",
  "(a)?( )?powell",
  "(m)?( )?pruitt",
  "(s)?( )?rice",
  "(t)?( )?richardson",
  "(c)?( )?sager",
  "(r)?( )?sherrell",
  "(b)?( )?smith",
  "(t)?( )?strong",
  "(m)?( )?tay(h)?lor",
  "(c)?( )?thom(a)?s",
  "(j)?( )?tumbrink",
  "(p)?( )?tyler",
  "(j)?( )?walker",
  "(m)?( )?ward",
  "(k)?( )?washington",
  "(b)?( )?woodland",
  "(v)?( )?woods",
  "ch[am]*bers",
  "ferguson",
  "funches",
  "griffin",
  "johnson",
  "judy",
  "lynum",
  "rimmer",
  "tidwell",
  "tyrone",
  "waler",
  "hokes",
  "eddie",
  "muhammad",
  "jn", #' judith norman
  "ma",
  "mc", "otis",
  "sh(?! e )" #' s howell
), "\\b") %>% 
  str_c(collapse = "|")
```


```{r}
resCode <- zNum %>%
  mutate(
    rc2 = str_extract_all(rs, vRC),
    person = str_extract_all(rs, vInspectors),
    rs = str_remove_all(rs, vRC),
    rs = str_remove_all(rs, vInspectors),
    rs = str_squish(rs),
    rs = na_if(rs, "")
  )

RC <- resCode %>% filter(!is.na(rs))
```


### Code Inspectors

There is frequent mention of code inspectors assigned to the case (per inspector james, walker, burton, etc).

To organize this data, I first identify mentions of certain people, including inspectors and tenants.

```{r}
vLV <- str_c("\\b", c("per", "by", "from", "(?<!due )to"), "\\b") %>% 
  str_c(collapse = "|")

vAdj <- str_c("\\b", c(
  "new", "old", "active", "prior", "final", "wrong", "incorrect", "insuf[fic]+ent", "inaccurate", "inadequate", "hazardous", "error", "all", "duplicate", 
  "no[tn]?", "[1-4](st|nd|rd|th)?", "one", "two", "three", "third", "second", "four", "dead", "further", "a[dit]+onal", "another", "also", "high",
  "hot", "derelict", "junky", "interior", "exterior", "(in)?valid", "void", "good", "great", "minor", "major", "adjacent", "under",
  "substandard", "rear", "still", "raw", "multiple", "st[ru]+ctural", "back", "front", "commer[cia]+l", "administrative",
  "overgrown", "official", "public", "broken", "bad", "around", "environmental(?! enforcement)", "inside", "large",
  "stagnant", "tall", "under", "behind", "excessive", "central", "original", "between", "same", "different", "black", "empty", "done", "cold", "green",
  "through(out)?", "rotten", "nice", "very", "above", "accessory", "other", "anti", "hard", "loose", "near", "across", "unsafe", "various", "below",
  "unknown", "missing", "pre", "well", "torn", "local", "loud"
  ), "\\b") %>% 
  str_c(collapse = "|")

vRoles <- str_c(
  "\\b",
  c(
    "in[sp]+([ector]+)?(?!ect(ion|ed))",
    "sup(ervisor)?(v)?", 
    "ten[nate]+t(ed)?",
    "occupant", "neighbor", "daughter", "son", "family", "community",
    "resident", "caregiver",
    "land(l)?ord",
    "(home)?owner(ship)?",
    "manage(r|ment)", "mgr",
    "company", "agent",
    "employee",
    "mgm(n)?t",
    "courts",
    "citizen", "people",
    "(grounds|health) dep(artmen)?t",
    "ground(s)? services",
    "condemnation",
    "(vector|animal|pest) control",
    "animal shelter",
    "sanitation",
    "solid waste",
    "environmental enforcement",
    "police", "ems",
    "(dep(artmen)?t)",
    "code enforcement",
    "(shelby )?county( (land bank|code))?",
    "city(of memphis)?",
    "maintenance", "contractor", "electrician", "a(r)?bor(t)?ist",
    "vend[eo]r",
    "(s)?he", "they", "their", "them", "m[rs]+", "noone", "my", "me", "we", "his", "her", "us",
    "nobody", "anyone", "man", "someone( else)?", "you", "i(?!(s| nsp))", "husband", "him",
    "mpd", "mlg( )?w", "individual",
    "memphis", "shelby", "u s coast guard",
    "mcsc", "mscs", "itd",
    "e[sz]ri",
    "csa", "wooddale", "foxhall", "fox hollow", "crestcore", "reedy", "revid", "barbara estate"
  ), "(( )?(s|es))?", "\\b"
) %>% str_c(collapse = "|")
  
vArticles <- str_c("\\b", c("(th|ht)e", "and", "a(?! c\\b)(n)?", "that", "this", "of", "where", "who", "with", "but", "or", "as", "fu(r)?thermore"), "\\b") %>% 
  str_c(collapse = "|")

```

This information is separated into a new column and the dataset is again tidied up.

```{r}
people <- RC %>% 
  mutate(
    LVppl = str_extract_all(rs, vLV),
    role = str_extract_all(rs, vRoles),
    adj = str_extract_all(rs, vAdj),
    rs = str_remove_all(rs, vAdj),
    rs = str_remove_all(rs, vArticles),
    rs = str_remove_all(rs, vRoles),
    rs = str_remove_all(rs, vLV),
    rs = str_squish(rs),
    rs = na_if(rs, "")
  )

ppl <- people %>% 
  filter(!is.na(rs))
```

Next, I created a list of names of code inspectors. I did this by filtering the `roles` column for mentions of "inspector", and manually building a list.

I then extracted inspector names into a separate column titled `person`.


Once again, filter out rows that are now empty and count the most common entries that remain in the resolution summary.

There are XXXXX rows that still contain information in the resolution summary.


### Status Updates

Rather than an inspector manually typing "at the time of inspection", there would ideally be a column for common updates to service requests. Using the above table, I have created an incomplete list of some of the most common updates.

```{r}
vAdv <- str_c(
  "\\b",
  c(
    "for",
    "already",
    "longer",
    "(?<!follow )up",
    "out", "new",
    "over", "into",
    "down", "full", "apparent",
    "off", "actual",
    "(in)?operable", "according",
    "please",
    "proper", "inadvertant",
    "previous", "former", "retaliatory",
    "casual", "everything",
    "adequate", "enough", "more", "any", "several", "only",
    "complete", "entire", "few", "nothing", "both", "none",
    "manual", "immediate", "some", "improper", "illegal",
    "total", "present", "possibl[le]*", "partial", "severe", "neat"
  ),
  "(ly)?",
  "\\b"
) %>%
  str_c(collapse = "|")

vProcess <- str_c(
  "\\b", "(re( )?)?",
  c(
    "case", "file", "npa", "court order", "\\bsr\\b",
    "notice", "court", "tax sale", "tire ordinance 5334",
    "(housing )?(code )?violation", "(?<!no )problem", 
    "civil( (matter|issue))", "hearing", "order",
    "(re)?inspection", "recheck", "apartment complex sweep", "investigation",
    "pic(ture)?", "photo", "repairs", "info(rmation)?", "assistance",
    "ph", "location", "junk", "door", "street", "truck", "cornice",
    "pro(p)?ert( )?(y|i)?", "(town)?house", "home", "structur(e|al)", "unit", 
    "apar(t)?ment", "address(?!ed)", "roof", "shed", "chatt[el]+", "sh(r)?ub(bery)?", "bush", "hedge",
    "grass", "vehicle", "car(d)?", "yard", "tree", "limb", "fenc(e|ing)", "branch", "leaves",
    "trash", "debris", "weed", "pool", "ticket", "(service )?request",
    "plumbing", "water", "leak", "light(ing)?", "stove", "toilet", "walkway",
    "a( )?c\\b", "air( condition(er|in))?", "window", "(re)?fri(d)?ge(rator)?",
    "trailer", "dumpster", "sidewalk", "garage", "furniture", "lot",
    "summon", "(out)?building", "ceiling", "attachment", "sewage",
    "electric(ity)?", "referral", "apt", "condos", "mo[lo]d(( )?like)?", "curb(side)?",
    "backyard", "tires", "furnace", "gas", "complex", "heat(er|ing)?", "driveway", "carport",
    "utilit(i|y)", "lawn", "fire", "rent", "residence", "bed( )?b[eu]g", "rodent", "rat", "mice", "pest", "roach",
    "number", "code", "l[ei]tigation", "invoice", "feces", "emergency", "mosquito", "spiders",
    "hvac", "gutter", "ramp", "proof", "couch", "faucet", "tub", "floor(ing)?", "it",
    "storage", "residential", "outside", "limit", "overgrowth", "parcel( id)?", "duplex",
    "tag", "fine", "gate", "sign", "office", "hanger", "bag", "basketball goal", "trampoline",
    "business", "customer", "wall", "garbage", "vegetation", "church", "triplex",
    "appliance", "camper", "alleyway", "dishwasher", "graffiti", "rooster", "chigger",
    "bathroom", "(tele)?phone", "porch", "evidence", "siding", "infestation",
    "pallet", "elevator", "concrete", "dog", "cabinet", "chicken", "hole",
    "pipe", "sink", "trim", "substance", "letter", "(bench )?warrant", "tarp",
    "project", "wood", "patio", "premise", "stair", "sweep", "facility", "stain",
    "odor", "smell", "school", "tax", "step", "side", "trap", "shower",
    "stump", "tire", "permit", "mildew", "condition", "district", "document",
    "auto", "shop", "plant", "bamboo", "occupancy", "item", "housing", "leasing",
    "area", "place", "process", "system", "lock", "socket", "class", "kitchen",
    "storm", "egress", "outlet", "program", "boat", "arrival", "glass", "generator",
    "job", "note", "handle", "paperwork", "sewer", "party", "cart", "breaker box", "citation",
    "weather", "hallway", "spore", "ipmc", "outdoor", "carpet", "knob", "voicemail",
    "restroom", "mistake", "involvement", "ground", "caller", "matter", "health", "safety", "construction", "test",
    "contract", "criteria", "judgement", "form", "hotel", "power", "vent(ilation)?", "vine",
    "status", "brick", "roofing", "room", "screen", "surface", "vinyl", "waste",
    "container", "service", "pothole", "pile", "survey", "legal hold", "ordinance",
    "proceeding", "noise", "bath", "box", "docket", "hazard", "mattress", "wire",
    "saw", "traffic", "acid", "bug powder", "situation", "action", "oracle", "dirt",
    "food", "vagrant", "dwelling", "washer", "dryer", "311", "site", "covid( 19)?", "cable",
    "litter", "roadside", "pesticide", "mask", "table", "task"
    ), "(( )?(es|s|(i?)al))?", "\\b"
) %>% str_c(collapse = "|")

vLV2 <- str_c("\\b", c("has(n t)?", "had", "have", "been", "was", "there", "what", "were", "will", "would", "is(n t)?", "(c|sh)ould(n t)?", "be", "because", "about", "being", "cannot", "do(n t)?", "did(n t)?", "does(n t)?", "are"), "\\b") %>% 
  str_c(collapse = "|")

vStatus <- str_c(
  "\\b", "(re( )?)?",
  c("ma[dk](e)?", "present", "ok", "add", "help",
  "s[lov]+e", "c[ao]m(e)?", "arrive", "engage", "instruct",
  "model", "stop", "gone",
  "clean(up)?", "f[eai]ll", "adopt", "cit(e)?", "dispose", "dilapidate",
  "repair", "r[ea]hab(b)?", "renovat(e)?", "destroy", "disable",
  "maintain", "abandon", "prune", "rent", "remain", "revise",
  "correct", "inform", "pend", "gain", "care", "explain",
  "just([if]+)?f(y|i)", "request",
  "fix", "touch", "bark", "post", "via", "forward", "refuse",
  "locat(e)?", "pass", "abat(e)?(ment)?", "accomplish", "acquire",
  "determine", "imperil", "caus(e)?", "go", "stay",
  "exterminate", "unclog", "satisf(y|i)", "decide",
  "burn", "attach", "abate", "turn", "void", "foreclose", "peel",
  "function", "enter", "list(?!\\b)", "flood",
  "clear", "confirm", "(a)?wait", "belong", "check", "reset", "alive", "bust",
  "(un)?able", "drain", "set", "provide",
  "walk", "tow", "suppl(y|i)",
  "t[ao]k(e)?", "avoid",
  "select", "treat", "place(?!\\b)", "lock(?!\\b)",
  "deliver", "smell",
  "eliminate", "cease", "block", "neglect",
  "replace",
  "g[eo]t", "look", "notice", "obstruct",
  "complete",
  "demo(lish)?", "print", "disinfect", "seal",
  "dismiss", "select",
  "paint", "spray", "dump",
  "fall", "pick", "(write|pick)up",
  "progress", "wear",
  "lean", "leak",
  "(?<!civil )issu(e)?",
  "start", "order",
  "boar(d)?", "secure",
  "creat(e)?",
  "need", "want", "meet",
  "provide", "includ(e)?",
  "cut", "mow",
  "left",
  "run(n)?",
  "give",
  "liv(e)?",
  "talk", "spoke", "state", "sa(id|y)", "mention",
  "see", "respond",
  "mov(e)?", "report",
  "own", "evict", "sold", "purchase", "sell",
  "work",
  "occup(i|y)",
  "vaca(n)?t",
  "found", "observe", "cover",
  "access", "assess", "show",
  "attempt",
  "a[d]+ress",
  "require",
  "cancel(l)?", "clos(e)?", "open",
  "exist",
  "response",
  "wr[oi][t]+(e)?", "sen[dt]", "email", "schedule", "train", 
  "mon[it]+or", "verif(y|i)", "approve",
  "outdate",
  "refe(r)+(?!al)",
  "entry",
  "comply",
  "damag(e)?", "concern", "receive",
  "hang",
  "us(e)?", "mail",
  "assign",
  "contact", "answer", "call",
  "submit(t)?",
  "install",
  "(in|non)?comp[lia]+n(ce|t)",
  "(?<!code )v[ioa]+l(a)?t(e|((i)?on))?",
  "park(?!\\b)",
  "trimm",
  "(un)?( )?c[op]+er[tai]+(v(i)?e)?",
  "damage",
  "restore",
  "inspect(?!ion)",
  "transfer", "allow", "f(i|ou)nd",
  "keep", "visit", "appear", "change", "update"
  ), 
  "(es|s|en|n|[elita]*ed|d|[lita]*ion(s)?|(ar)?ing)?", 
  "\\b"
  ) %>% str_c(collapse = "|")

vEvent <- str_c("\\b", c(
  "at.*time.*of", "on", "f( )?u", "if", "at", "upon", "time", "next", "due to",
  "((follow up|later|next) )?date", "(with)?in", "now", "follow up", "continue", "currently",
  "during", "\\d{1,3} day(s)?", "weather permi[t]+ing", "current", "\\d{1,2}( )?h(ou)?r(s)?",
  "month(s( ago)?)?", "week(end|s)?", "never", "before", "recently", "until", "again", "after",
  "yesterday", "since", "april", "january", "friday", "saturday", "monday", "future", "today", "spring( 2021)?", "year",
  "morning", "once", "while", "when", "subsequent", "2019", "soon", "earlier", "time"
  ), "\\b") %>% 
  str_c(collapse = "|")
```

```{r}
subject <- ppl %>% 
  mutate(
    LV2 = str_extract_all(rs, vLV2),
    event = str_extract_all(rs, vEvent),
    status = str_extract_all(rs, vStatus),
    process = str_extract_all(rs, vProcess),
    rs = str_remove_all(rs, vAdv),
    rs = str_remove_all(rs, vLV2),
    rs = str_remove_all(rs, vEvent),
    rs = str_remove_all(rs, vStatus),
    rs = str_remove_all(rs, vProcess),
    rs = str_squish(rs),
    rs = na_if(rs, "")
  )

sbj <- subject %>% 
  filter(!is.na(rs))

sbj %>% count(rs) %>% arrange(desc(n))
```


