---
title: "Census Housing Variables"
description: |
  A visual tree of Census variables related to housing.
author:
  - name: Sarah Johnson
date: 2022-06-18
categories:
  - census
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 4
    code_download: true
repository_url: https://github.com/sj-io/sj-distill
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The Census releases [a lot of information](https://api.census.gov/data/2020/acs/acs5/variables.html). I wanted a way to visualize Census variables related to housing, so I made a tree using R and the terminal. The following tree shows 2020 Census variables that start with "B25", the housing-related group, excluding aggregates, medians, or race subvariables.

# Census variable tree

``` zsh
.
├── AGE-OF-HOUSEHOLDER
│   ├── GROSS-RENT-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   │   └── 25072
│   ├── MEALS-INCLUDED-IN-RENT
│   │   └── 25055
│   └── SELECTED-MONTHLY-OWNER-COSTS-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│       └── 25093
├── BEDROOMS
│   ├── 25041
│   └── GROSS-RENT
│       └── 25068
├── CONTRACT-RENT
│   └── 25056
├── GROSS-RENT
│   └── 25063
├── GROSS-RENT-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   └── 25070
├── HOUSE-HEATING-FUEL
│   └── 25040
├── HOUSEHOLD-INCOME
│   ├── GROSS-RENT-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   │   └── 25074
│   └── SELECTED-MONTHLY-OWNER-COSTS-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│       └── 25095
├── HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS-IN-2020-INFLATION-ADJUSTED-DOLLARS
│   ├── GROSS-RENT
│   │   └── 25122
│   └── VALUE
│       └── 25121
├── HOUSING-UNITS
│   └── 25001
├── INCLUSION-OF-UTILITIES-IN-RENT
│   └── 25069
├── KITCHEN-FACILITIES
│   └── MEALS-INCLUDED-IN-RENT
│       └── 25054
├── KITCHEN-FACILITIES-FOR-ALL-HOUSING-UNITS
│   └── 25051
├── KITCHEN-FACILITIES-FOR-OCCUPIED-HOUSING-UNITS
│   └── 25052
├── MONTHLY-HOUSING-COSTS
│   └── 25104
├── MORTGAGE-STATUS
│   ├── 25081
│   ├── AGE-OF-HOUSEHOLDER
│   │   └── 25027
│   ├── HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS-IN-2020-INFLATION-ADJUSTED-DOLLARS
│   │   └── 25098
│   ├── MONTHLY-HOUSING-COSTS-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   │   └── 25101
│   ├── RATIO-OF-VALUE-TO-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   │   └── 25100
│   ├── REAL-ESTATE-TAXES-PAID
│   │   └── 25102
│   ├── SELECTED-MONTHLY-OWNER-COSTS-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   │   └── 25091
│   └── VALUE
│       └── 25096
├── MORTGAGE-STATUS-AND-SELECTED-MONTHLY-OWNER-COSTS
│   └── 25087
├── OCCUPANCY-STATUS
│   └── 25002
├── PLUMBING-FACILITIES
│   └── OCCUPANTS-PER-ROOM
│       └── YEAR-STRUCTURE-BUILT
│           └── 25050
├── PLUMBING-FACILITIES-FOR-ALL-HOUSING-UNITS
│   └── 25047
├── PLUMBING-FACILITIES-FOR-OCCUPIED-HOUSING-UNITS
│   └── 25048
├── PRICE-ASKED
│   └── 25085
├── RACE-OF-HOUSEHOLDER
│   └── 25006
├── RENT-ASKED
│   └── 25061
├── ROOMS
│   └── 25017
├── SELECTED-MONTHLY-OWNER-COSTS
│   └── 25094
├── TENURE
│   ├── 25003
│   ├── AGE-OF-HOUSEHOLDER
│   │   ├── 25007
│   │   ├── OCCUPANTS-PER-ROOM
│   │   │   └── 25015
│   │   ├── UNITS-IN-STRUCTURE
│   │   │   └── 25125
│   │   ├── YEAR-HOUSEHOLDER-MOVED-INTO-UNIT
│   │   │   └── 25128
│   │   └── YEAR-STRUCTURE-BUILT
│   │       └── 25126
│   ├── BEDROOMS
│   │   └── 25042
│   ├── EDUCATIONAL-ATTAINMENT-OF-HOUSEHOLDER
│   │   └── 25013
│   ├── FAMILIES-AND-PRESENCE-OF-OWN-CHILDREN
│   │   └── 25012
│   ├── HOUSE-HEATING-FUEL
│   │   └── 25117
│   ├── HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS-IN-2020-INFLATION-ADJUSTED-DOLLARS
│   │   └── 25118
│   ├── HOUSEHOLD-SIZE
│   │   ├── 25009
│   │   ├── AGE-OF-HOUSEHOLDER
│   │   │   └── 25116
│   │   └── UNITS-IN-STRUCTURE
│   │       └── 25124
│   ├── HOUSEHOLD-TYPE-AND-PRESENCE-AND-AGE-OF-OWN-CHILDREN
│   │   └── 25115
│   ├── HOUSEHOLD-TYPE-INCLUDING-LIVING-ALONE-AND-AGE-OF-HOUSEHOLDER
│   │   └── 25011
│   ├── HOUSING-COSTS-AS-A-PERCENTAGE-OF-HOUSEHOLD-INCOME-IN-THE-PAST-12-MONTHS
│   │   └── 25106
│   ├── KITCHEN-FACILITIES
│   │   └── 25053
│   ├── OCCUPANTS-PER-ROOM
│   │   └── 25014
│   ├── PLUMBING-FACILITIES
│   │   ├── 25049
│   │   └── OCCUPANTS-PER-ROOM
│   │       └── 25016
│   ├── ROOMS
│   │   └── 25020
│   ├── SELECTED-PHYSICAL-AND-FINANCIAL-CONDITIONS
│   │   └── 25123
│   ├── TELEPHONE-SERVICE-AVAILABLE
│   │   └── AGE-OF-HOUSEHOLDER
│   │       └── 25043
│   ├── UNITS-IN-STRUCTURE
│   │   └── 25032
│   ├── VEHICLES-AVAILABLE
│   │   ├── 25044
│   │   └── AGE-OF-HOUSEHOLDER
│   │       └── 25045
│   ├── YEAR-HOUSEHOLDER-MOVED-INTO-UNIT
│   │   ├── 25038
│   │   └── UNITS-IN-STRUCTURE
│   │       └── 25129
│   └── YEAR-STRUCTURE-BUILT
│       ├── 25036
│       └── UNITS-IN-STRUCTURE
│           └── 25127
├── TOTAL-POPULATION-IN-OCCUPIED-HOUSING-UNITS
│   └── TENURE
│       ├── 25008
│       ├── UNITS-IN-STRUCTURE
│       │   └── 25033
│       └── YEAR-HOUSEHOLDER-MOVED-INTO-UNIT
│           └── 25026
├── UNITS-IN-STRUCTURE
│   └── 25024
├── VACANCY-STATUS
│   └── 25004
├── VACANT-CURRENT-RESIDENCE-ELSEWHERE
│   └── 25005
├── VALUE
│   └── 25075
└── YEAR-STRUCTURE-BUILT
    └── 25034
    
160 directories, 0 files
```

# The Code

## R

First, load the tidyverse and tidycensus packages.

```{r}
library(tidyverse)
library(tidycensus)
```

Use the `load_variables()` function to get Census variables.

```{r}
v20 <- load_variables(2020, "acs5")
```

Narrow for housing-related variables, which start with "B25".

```{r}
hsg <- v20 %>% 
  filter(str_detect(name, "^B25"))

head(hsg)
```

There are 2,208 variables related to housing, so I narrowed down the data using `filter()`. I used `mutate()` to change variable names into a folder-friendly format. Then, `unite()` was used to join the variable name with the variable, separated by a slash.

```{r}
h1 <- hsg %>% 
  filter(str_detect(name, "_001$"),           #' only top-level variables
         !str_detect(name, "[:alpha:]_"),     #' omit race variables
         str_detect(label, "Estimate!!Total") #' omit aggregates, medians, etc.
         ) %>% 
  select(-label) %>%
  mutate(concept = str_replace_all(concept, "-(?=[:alpha:])", " "),
         concept = str_remove_all(concept, "[:punct:]"),
         concept = str_replace_all(concept, " BY ", "/"),
         name = str_remove(name, "_001"),
         name = str_remove_all(name, "^B")
         ) %>% 
  unite(col = "fileStr", concept, name, sep = "/", na.rm = TRUE) %>% 
  mutate(fileStr = str_replace_all(fileStr, "\\s", "-"),
         fileStr = str_replace_all(fileStr, "-{2,}", "-"))

head(h1)
```

Turn the column into a list of values.

```{r}
vH1 <- h1$fileStr
```

Use `write_lines()` to convert the list into a .txt file.

```{r}
write_lines(vH1, file = "~/Documents/H1.txt")
```

The output is a text file with 76 lines, which we will convert to a directory.

## Terminal

In the terminal, make a new directory to house your Census tree.

``` zsh
mkdir ~/Documents/census
cd Documents/census
```

Convert the text file to a file directory using the following code.

``` zsh
xargs -I {} mkdir -p "{}" < "../H1.txt"
```

Finally, create a tree of the file directory. Note: requires the [tree](Turn the column into a list of values.) package.

``` zsh
tree
```

The output is the tree at the top of this page.

You can search for certain words in the terminal using `find`.

``` zsh
find . -name "AGE*"
./AGE-OF-HOUSEHOLDER
./TENURE/AGE-OF-HOUSEHOLDER
./TENURE/VEHICLES-AVAILABLE/AGE-OF-HOUSEHOLDER
./TENURE/TELEPHONE-SERVICE-AVAILABLE/AGE-OF-HOUSEHOLDER
./TENURE/HOUSEHOLD-SIZE/AGE-OF-HOUSEHOLDER
./MORTGAGE-STATUS/AGE-OF-HOUSEHOLDER
```
